---
title: "I/O Performance Analysis"
subtitle: "OTF2 Trace Visualization"
execute:
  echo: false
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
---

```{r, include=FALSE}
library(jsonlite)

results <- fromJSON("results.json")

ticks_per_second <- results$TimerResolution

format_bytes <- function(bytes) {
  if (is.na(bytes)) return("0 B")
  bytes <- as.numeric(bytes)
  if (bytes >= 1e15) return(sprintf("%.2f PB", bytes / 1e15))
  if (bytes >= 1e12) return(sprintf("%.2f TB", bytes / 1e12))
  if (bytes >= 1e9)  return(sprintf("%.2f GB", bytes / 1e9))
  if (bytes >= 1e6)  return(sprintf("%.2f MB", bytes / 1e6))
  if (bytes >= 1e3)  return(sprintf("%.2f KB", bytes / 1e3))
  return(sprintf("%d B", bytes))
}

format_time <- function(ticks) {
  if (is.na(ticks)) return("0 s")
  ticks <- as.numeric(ticks)
  seconds <- ticks / ticks_per_second
  if (seconds >= 3600) return(sprintf("%.2f h", seconds / 3600))
  if (seconds >= 60)   return(sprintf("%.2f min", seconds / 60))
  if (seconds >= 1)    return(sprintf("%.2f s", seconds))
  if (seconds >= 1e-3) return(sprintf("%.2f ms", seconds * 1e3))
  if (seconds >= 1e-6) return(sprintf("%.2f µs", seconds * 1e6))
  return(sprintf("%.2f ns", seconds * 1e9))
}
```

```{r, include=FALSE}
# Compute summary metrics
io_ops <- results$IOOperations
total_bytes <- sum(sapply(io_ops, function(x) ifelse(is.null(x$Bytes), 0, x$Bytes)), na.rm = TRUE)
total_time <- sum(sapply(io_ops, function(x) {
  meta <- ifelse(is.null(x$MetaOperationTime), 0, x$MetaOperationTime)
  trans <- ifelse(is.null(x$TransferOperationTime), 0, x$TransferOperationTime)
  (meta + trans) / ticks_per_second
}), na.rm = TRUE)
total_ops <- sum(sapply(io_ops, function(x) ifelse(is.null(x$Count), 0, x$Count)), na.rm = TRUE)
num_files <- nrow(results$Files)
```

## Overview

**Total I/O:** `r format_bytes(total_bytes)` | **Total I/O Time:** `r format_time(total_time * ticks_per_second)` | **I/O Operations:** `r total_ops` | **Files:** `r num_files`

**Trace:** `r results$Trace$FileName` | **Processes:** `r results$ProcessCount` | **Timer Resolution:** `r sprintf("%.2f ms", results$TimerResolution / 1e6)`

## I/O by Paradigm

```{r, fig.width=10, fig.height=5}
# Prepare paradigm data
paradigms <- names(io_ops)
bytes_data <- sapply(io_ops, function(x) ifelse(is.null(x$Bytes), 0, x$Bytes))
time_data <- sapply(io_ops, function(x) {
  meta <- ifelse(is.null(x$MetaOperationTime), 0, x$MetaOperationTime)
  trans <- ifelse(is.null(x$TransferOperationTime), 0, x$TransferOperationTime)
  (meta + trans) / ticks_per_second
})

par_df <- data.frame(
  Paradigm = paradigms,
  Bytes = bytes_data,
  Time = time_data,
  stringsAsFactors = FALSE
)

# Plot bytes by paradigm
par_df$Paradigm <- factor(par_df$Paradigm, levels = par_df$Paradigm[order(par_df$Bytes)])
barplot(par_df$Bytes, names.arg = par_df$Paradigm, horiz = TRUE, 
        xlab = "Bytes", main = "Total I/O by Paradigm",
        col = rainbow(length(par_df$Paradigm)))
```

```{r, fig.width=10, fig.height=5}
# Plot time by paradigm
par_df$Paradigm <- factor(par_df$Paradigm, levels = par_df$Paradigm[order(par_df$Time)])
barplot(par_df$Time, names.arg = par_df$Paradigm, horiz = TRUE,
        xlab = "Time", main = "I/O Time by Paradigm",
        col = rainbow(length(par_df$Paradigm)))
```

## Local Access Patterns

```{r, fig.width=10, fig.height=6}
# Aggregate access pattern data from data.frame
ap_ticks <- results$Files$`Ticks spent per Access Pattern`
ap_bytes <- results$Files$`I/O sizes per Access Pattern`

# Sum across all files
total_ticks <- colSums(ap_ticks, na.rm = TRUE)
total_bytes_ap <- colSums(ap_bytes, na.rm = TRUE)

patterns <- gsub("AccessPattern::", "", names(total_ticks))
ap_df <- data.frame(
  Pattern = patterns,
  Ticks = as.numeric(total_ticks),
  Bytes = as.numeric(total_bytes_ap),
  Time = as.numeric(total_ticks) / ticks_per_second,
  stringsAsFactors = FALSE
)
ap_df <- ap_df[ap_df$Bytes > 0 | ap_df$Time > 0, ]

# Plot access pattern distribution
if (nrow(ap_df) > 0) {
  ap_df$Pattern <- factor(ap_df$Pattern, levels = ap_df$Pattern[order(ap_df$Time)])
  barplot(ap_df$Time, names.arg = ap_df$Pattern, horiz = TRUE,
          xlab = "Time", main = "I/O Time by Access Pattern",
          col = rainbow(length(ap_df$Pattern)))
}
```

```{r, fig.width=10, fig.height=6}
# Plot bytes by access pattern
if (nrow(ap_df) > 0) {
  ap_df$Pattern <- factor(ap_df$Pattern, levels = ap_df$Pattern[order(ap_df$Bytes)])
  barplot(ap_df$Bytes, names.arg = ap_df$Pattern, horiz = TRUE,
          xlab = "Bytes", main = "I/O Size by Access Pattern",
          col = rainbow(length(ap_df$Pattern)))
}
```

## Global Access Pattern Analysis

```{r, fig.width=10, fig.height=5}
# Aggregate by number of locations
loc_agg <- aggregate(
  cbind(`#Bytes read`, `#Bytes write`) ~ `Nr accesses from different locations`, 
  data = results$Files, 
  FUN = sum, 
  na.rm = TRUE
)
loc_agg$TotalBytes <- loc_agg$`#Bytes read` + loc_agg$`#Bytes write`
loc_agg$Label <- paste0(loc_agg$`Nr accesses from different locations`, 
                        ifelse(loc_agg$`Nr accesses from different locations` == 1, " rank", " ranks"))
loc_agg <- loc_agg[order(-loc_agg$TotalBytes), ]

# Plot by location count
barplot(loc_agg$TotalBytes, names.arg = loc_agg$Label, horiz = TRUE,
        xlab = "Bytes", main = "I/O by Number of Accessing Ranks",
        col = "#3498db")
```

## Top Files

```{r}
# Top files by I/O
top_files <- results$Files
top_files$Total <- top_files$`#Bytes read` + top_files$`#Bytes write`
top_files$Time <- (top_files$`Ticks spent on I/O` + top_files$`Ticks spent on Metadata`) / ticks_per_second
top_files <- top_files[order(-top_files$Total), ][1:10, ]
top_files$FileName <- gsub(".*/", "", top_files$FileName)

# Print as markdown table
cat("| File | Read | Write | Time | Paradigm |\n")
cat("|------|------|-------|------|----------|\n")
for (i in 1:nrow(top_files)) {
  paradigm <- paste(top_files$IoParadigm[[i]], collapse = ", ")
  cat(sprintf("| %s | %s | %s | %s | %s |\n",
              top_files$FileName[i],
              format_bytes(top_files$`#Bytes read`[i]),
              format_bytes(top_files$`#Bytes write`[i]),
              format_time(top_files$Time[i] * ticks_per_second),
              paradigm))
}
```

---

*Generated by OTF-Profiler | Quarto Visualization*
