---
title: "I/O Performance Analysis"
subtitle: "OTF2 Trace Visualization"
execute:
  echo: false
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
---

```{r, include=FALSE}
library(jsonlite)

results <- fromJSON("results.json")

ticks_per_second <- results$TimerResolution

format_bytes <- function(bytes) {
  if (is.na(bytes)) return("0 B")
  bytes <- as.numeric(bytes)
  if (bytes >= 1e15) return(sprintf("%.2f PB", bytes / 1e15))
  if (bytes >= 1e12) return(sprintf("%.2f TB", bytes / 1e12))
  if (bytes >= 1e9)  return(sprintf("%.2f GB", bytes / 1e9))
  if (bytes >= 1e6)  return(sprintf("%.2f MB", bytes / 1e6))
  if (bytes >= 1e3)  return(sprintf("%.2f KB", bytes / 1e3))
  return(sprintf("%d B", bytes))
}

format_time <- function(ticks) {
  if (is.na(ticks)) return("0 s")
  ticks <- as.numeric(ticks)
  seconds <- ticks / ticks_per_second
  if (seconds >= 3600) return(sprintf("%.2f h", seconds / 3600))
  if (seconds >= 60)   return(sprintf("%.2f min", seconds / 60))
  if (seconds >= 1)    return(sprintf("%.2f s", seconds))
  if (seconds >= 1e-3) return(sprintf("%.2f ms", seconds * 1e3))
  if (seconds >= 1e-6) return(sprintf("%.2f µs", seconds * 1e6))
  return(sprintf("%.2f ns", seconds * 1e9))
}
```

```{r, include=FALSE}
# Compute summary metrics
io_ops <- results$IOOperations
total_bytes <- sum(sapply(io_ops, function(x) ifelse(is.null(x$Bytes), 0, x$Bytes)), na.rm = TRUE)
total_time <- sum(sapply(io_ops, function(x) {
  meta <- ifelse(is.null(x$MetaOperationTime), 0, x$MetaOperationTime)
  trans <- ifelse(is.null(x$TransferOperationTime), 0, x$TransferOperationTime)
  (meta + trans) / ticks_per_second
}), na.rm = TRUE)
total_ops <- sum(sapply(io_ops, function(x) ifelse(is.null(x$Count), 0, x$Count)), na.rm = TRUE)
num_files <- nrow(results$Files)
```

## Overview

**Total I/O:** `r format_bytes(total_bytes)` | **Total I/O Time:** `r format_time(total_time * ticks_per_second)` | **I/O Operations:** `r total_ops` | **Files:** `r num_files`

**Trace:** `r results$Trace$FileName` | **Processes:** `r results$ProcessCount` | **Timer Resolution:** `r sprintf("%.2f ms", results$TimerResolution / 1e6)`

## I/O by Paradigm

```{r, fig.width=12, fig.height=5}
# Prepare paradigm data - all three metrics
paradigms <- names(io_ops)
bytes_data <- sapply(io_ops, function(x) ifelse(is.null(x$Bytes), 0, x$Bytes))
meta_data <- sapply(io_ops, function(x) ifelse(is.null(x$MetaOperationTime), 0, x$MetaOperationTime) / ticks_per_second)
transfer_data <- sapply(io_ops, function(x) ifelse(is.null(x$TransferOperationTime), 0, x$TransferOperationTime) / ticks_per_second)

par_df <- data.frame(
  Paradigm = paradigms,
  Bytes = bytes_data,
  MetaTime = meta_data,
  TransferTime = transfer_data,
  stringsAsFactors = FALSE
)

# Plot all three metrics side by side
par(mfrow = c(1, 3), mar = c(5, 8, 4, 2))

# Bytes
par_df <- par_df[order(par_df$Bytes, decreasing = TRUE), ]
barplot(par_df$Bytes, names.arg = par_df$Paradigm, horiz = TRUE, 
        xlab = "Bytes", main = "(a) I/O Size per Paradigm",
        col = rainbow(length(par_df$Paradigm)))

# Transfer Time  
par_df <- par_df[order(par_df$TransferTime, decreasing = TRUE), ]
barplot(par_df$TransferTime, names.arg = par_df$Paradigm, horiz = TRUE,
        xlab = "Time (s)", main = "(b) Transfer Time per Paradigm",
        col = rainbow(length(par_df$Paradigm)))

# Meta Time
par_df <- par_df[order(par_df$MetaTime, decreasing = TRUE), ]
barplot(par_df$MetaTime, names.arg = par_df$Paradigm, horiz = TRUE,
        xlab = "Time (s)", main = "(c) Meta I/O Time per Paradigm",
        col = rainbow(length(par_df$Paradigm)))

par(mfrow = c(1, 1))
```

## Local Access Patterns

*Analysis of per-file I/O access patterns (CONTIGUOUS, STRIDED, RANDOM, NONE)*

```{r, fig.width=12, fig.height=6}
# Aggregate access pattern data - both ticks and bytes
files <- results$Files

# Get access pattern columns
ap_ticks <- files$`Ticks spent per Access Pattern`
ap_bytes <- files$`I/O sizes per Access Pattern`

# Sum across all files for each pattern
patterns <- names(ap_ticks)
total_ticks <- colSums(ap_ticks, na.rm = TRUE)
total_bytes_ap <- colSums(ap_bytes, na.rm = TRUE)

pattern_names <- gsub("AccessPattern::", "", patterns)

ap_df <- data.frame(
  Pattern = pattern_names,
  Ticks = as.numeric(total_ticks),
  Bytes = as.numeric(total_bytes_ap),
  Time = as.numeric(total_ticks) / ticks_per_second,
  stringsAsFactors = FALSE
)
ap_df <- ap_df[ap_df$Bytes > 0 | ap_df$Time > 0, ]

# Plot both Time and Bytes side by side
par(mfrow = c(1, 2), mar = c(5, 10, 4, 2))

# Time by pattern
ap_df <- ap_df[order(ap_df$Time, decreasing = TRUE), ]
barplot(ap_df$Time, names.arg = ap_df$Pattern, horiz = TRUE,
        xlab = "Time (seconds)", main = "I/O Time by Access Pattern",
        col = rainbow(nrow(ap_df)))

# Bytes by pattern  
ap_df <- ap_df[order(ap_df$Bytes, decreasing = TRUE), ]
barplot(ap_df$Bytes, names.arg = ap_df$Pattern, horiz = TRUE,
        xlab = "Bytes", main = "I/O Size by Access Pattern",
        col = rainbow(nrow(ap_df)))

par(mfrow = c(1, 1))
```

### Per-File Access Pattern Breakdown

```{r, fig.width=12, fig.height=8}
# Create per-file access pattern table
file_names <- gsub(".*/", "", files$FileName)
bytes_read <- as.numeric(files$`#Bytes read`)
bytes_write <- as.numeric(files$`#Bytes write`)
ap_ticks <- files$`Ticks spent per Access Pattern`
ap_bytes <- files$`I/O sizes per Access Pattern`

# Filter to files with actual I/O
has_io <- (bytes_read + bytes_write) > 0

# Print as detailed table
cat("\n### Detailed Per-File Access Patterns\n\n")
cat("| File | Read | Write |")
cat(" CONTIGUOUS | STRIDED | RANDOM | NONE |")
cat(" CONG_Ticks | STRD_Ticks | RAND_Ticks | NONE_Ticks |\n")
cat("|--------|-----------|------------|", rep("--------|", 8), "\n")

for (i in 1:nrow(files)) {
  if (!has_io[i] || nchar(file_names[i]) == 0) next
  
  read_b <- format_bytes(bytes_read[i])
  write_b <- format_bytes(bytes_write[i])
  
  # Get pattern bytes
  cont_b <- format_bytes(ap_bytes[i, ]$`AccessPattern::CONTIGUOUS`)
  strd_b <- format_bytes(ap_bytes[i, ]$`AccessPattern::STRIDED`)
  rand_b <- format_bytes(ap_bytes[i, ]$`AccessPattern::RANDOM`)
  none_b <- format_bytes(ap_bytes[i, ]$`AccessPattern::NONE`)
  
  # Get pattern ticks
  cont_t <- ap_ticks[i, ]$`AccessPattern::CONTIGUOUS`
  strd_t <- ap_ticks[i, ]$`AccessPattern::STRIDED`
  rand_t <- ap_ticks[i, ]$`AccessPattern::RANDOM`
  none_t <- ap_ticks[i, ]$`AccessPattern::NONE`
  
  cat(sprintf("| %s | %s | %s | %s | %s | %s | %s | %d | %d | %d | %d |\n",
              substr(file_names[i], 1, 30), read_b, write_b,
              cont_b, strd_b, rand_b, none_b,
              cont_t, strd_t, rand_t, none_t))
}
```

## Global Access Pattern Analysis

*How I/O patterns vary based on number of MPI ranks accessing each file*

```{r, fig.width=10, fig.height=5}
# Aggregate by number of locations - also track patterns
files <- results$Files
loc_data <- list()

for (i in 1:nrow(files)) {
  nr_locs <- files$`Nr accesses from different locations`[i]
  if (nr_locs == 0) next
  
  bytes <- files$`#Bytes read`[i] + files$`#Bytes write`[i]
  ticks <- files$`Ticks spent on I/O`[i] + files$`Ticks spent on Metadata`[i]
  
  # Get patterns
  pat_bytes <- files$`I/O sizes per Access Pattern`[i, ]
  pat_ticks <- files$`Ticks spent per Access Pattern`[i, ]
  
  if (is.null(loc_data[[as.character(nr_locs)]])) {
    loc_data[[as.character(nr_locs)]] <- list(
      bytes = 0, files = 0, ticks = 0,
      cont_bytes = 0, strd_bytes = 0, rand_bytes = 0, none_bytes = 0,
      cont_ticks = 0, strd_ticks = 0, rand_ticks = 0, none_ticks = 0
    )
  }
  
  loc <- loc_data[[as.character(nr_locs)]]
  loc$bytes <- loc$bytes + bytes
  loc$files <- loc$files + 1
  loc$ticks <- loc$ticks + ticks
  loc$cont_bytes <- loc$cont_bytes + (pat_bytes$`AccessPattern::CONTIGUOUS` %||% 0)
  loc$strd_bytes <- loc$strd_bytes + (pat_bytes$`AccessPattern::STRIDED` %||% 0)
  loc$rand_bytes <- loc$rand_bytes + (pat_bytes$`AccessPattern::RANDOM` %||% 0)
  loc$none_bytes <- loc$none_bytes + (pat_bytes$`AccessPattern::NONE` %||% 0)
  loc$cont_ticks <- loc$cont_ticks + (pat_ticks$`AccessPattern::CONTIGUOUS` %||% 0)
  loc$strd_ticks <- loc$strd_ticks + (pat_ticks$`AccessPattern::STRIDED` %||% 0)
  loc$rand_ticks <- loc$rand_ticks + (pat_ticks$`AccessPattern::RANDOM` %||% 0)
  loc$none_ticks <- loc$none_ticks + (pat_ticks$`AccessPattern::NONE` %||% 0)
  loc_data[[as.character(nr_locs)]] <- loc
}

loc_df <- data.frame(
  Locations = as.integer(names(loc_data)),
  Bytes = sapply(loc_data, function(x) x$bytes),
  Files = sapply(loc_data, function(x) x$files),
  stringsAsFactors = FALSE
)
loc_df <- loc_df[order(-loc_df$Bytes), ]
loc_df$Label <- paste0(loc_df$Locations, ifelse(loc_df$Locations == 1, " rank", " ranks"))

# Plot by location count
par(mar = c(5, 10, 4, 2))
barplot(loc_df$Bytes, names.arg = loc_df$Label, horiz = TRUE,
        xlab = "Bytes", main = "I/O by Number of Accessing Ranks",
        col = "#3498db")
par(mar = c(5, 4, 4, 2))
```

### Access Patterns by Rank Count

```{r, fig.width=12, fig.height=5}
# Show pattern breakdown by rank count
loc_pattern_df <- data.frame(
  Label = loc_df$Label,
  CONTIGUOUS = sapply(loc_data, function(x) x$cont_bytes),
  STRIDED = sapply(loc_data, function(x) x$strd_bytes),
  RANDOM = sapply(loc_data, function(x) x$rand_bytes),
  NONE = sapply(loc_data, function(x) x$none_bytes)
)

# Stacked bar chart
par(mar = c(5, 10, 4, 2))
barplot(t(loc_pattern_df[, -1]), names.arg = loc_pattern_df$Label, horiz = TRUE,
        col = c("#2ecc71", "#3498db", "#e74c3c", "#95a5a6"),
        xlab = "Bytes", main = "Access Pattern Distribution by Rank Count",
        legend = TRUE)
par(mar = c(5, 4, 4, 2))
```

## Top Files

### Top 10 by I/O Size

```{r}
top_files <- results$Files
top_files$Total <- top_files$`#Bytes read` + top_files$`#Bytes write`
top_files$Time <- (top_files$`Ticks spent on I/O` + top_files$`Ticks spent on Metadata`) / ticks_per_second
top_files <- top_files[order(-top_files$Total), ][1:10, ]
top_files$FileName <- gsub(".*/", "", top_files$FileName)

cat("| File | Read | Write | Time | Paradigm |\n")
cat("|------|------|-------|------|----------|\n")
for (i in 1:nrow(top_files)) {
  paradigm <- paste(top_files$IoParadigm[[i]], collapse = ", ")
  cat(sprintf("| %s | %s | %s | %s | %s |\n",
              top_files$FileName[i],
              format_bytes(top_files$`#Bytes read`[i]),
              format_bytes(top_files$`#Bytes write`[i]),
              format_time(top_files$Time[i] * ticks_per_second),
              paradigm))
}
```

### Top 10 by I/O Time

```{r}
top_time <- results$Files
top_time$Time <- (top_time$`Ticks spent on I/O` + top_time$`Ticks spent on Metadata`) / ticks_per_second
top_time <- top_time[order(-top_time$Time), ][1:10, ]
top_time$FileName <- gsub(".*/", "", top_time$FileName)

cat("\n| File | I/O Time | Read | Write |\n")
cat("|------|----------|------|-------|\n")
for (i in 1:nrow(top_time)) {
  cat(sprintf("| %s | %s | %s | %s |\n",
              top_time$FileName[i],
              format_time(top_time$Time[i] * ticks_per_second),
              format_bytes(top_time$`#Bytes read`[i]),
              format_bytes(top_time$`#Bytes write`[i])))
}
```

---

*Generated by OTF-Profiler | Quarto Visualization*
